Supporting Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-28 Ben_Manuscripts/structure_paper/Draft.tex
Updates to the introduction

  %MRS13: I tweaked this a bit.  Probably could add another sentence to add some of the other important findings.
  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular, we compare
  simulated X-ray diffraction (XRD) patterns with experimental XRD to quantify and understand 
  differences between simulation and experiment.
  We find that pores are
  likely composed of 5 columns of stacked LLC monomers which surround each
  hydrophilic core.  We also find that some of the structure previously attributed to 
  monomer tail tilt is likely instead due to ordered tail packing.  
  Although this system has been reported as dry, we show that
  small amounts of water are necessary to fully reproduce the experimental XRD
  pattern due to asymmetries introduced by hydrogen bonds between monomer head
  groups and water molecules. We explore the composition and structure of the
  nanopores and reveal that there exists a composition gradient rather than a
  abrupt partition between the hydrophilic and hydrophobic regions. A caveat is
  that the time scales of the dynamics are extremely long for this system, resulting in 
  simulated structures that appear too ordered, 
  %MRS13: I left out starting from alternative structures since we didn't actually do it.
  %requiring starting from a range of alternative initial configurations and
  and requiring careful examination of the metastable states observed for any conclusions. The clearer picture of
  the nanoscopic structure of these membranes provided in this study will enable
  a better understanding of the mechanisms of small molecule transport within
  these nanopores.
  \end{abstract}

@@

  More highly selective membranes would be extremely useful for the recovery of
  dissolved species in complex aqueous and organic solutions. For example,
  selective membranes can further progress forward osmosis processes by reducing
  the cost of separating potable water from draw
  solutions~\cite{mccutcheon_novel_2005}. Additionally, flowback water (FW)
  produced during hydraulic fracturing is a complex wastewater full of
  potentially valuable dissolved organic compounds such as acetate that can be
  recovered using highly selective membranes \cite{dischinger_application_2017}.
  Finally, a significant amount of nitrogen and phosphorus from fertilizer is
  found in wastewater. Selective membranes offer a route towards low energy
  nutrient recovery from these wastewaters which can help sustain fertilizer
  production and subsequently food production into the future
  \cite{xie_membrane-based_2016}.
  % as we near peak production of minable phosphorus rock .

@@

% MRS6: I would keep a paragraph about here how current membranes DON'T
% work for this purpose. But one paragraph is enough.
% BJC5: added back below paragraph.  
% MRS8: ``because they are unstructured'' lacks detail.  NF has some structure, since there are irregular pores. Check some other reviews and see the points they cover?
  Current commercial RO and NF membranes suffer limitations to their selectivity
  because they are unstructured. Although scalable, their fabrication involves the 
  spontaneous assembly of polymers into disordered structures which offer separation 
  pathways that are tortuous and polydisperse in size. This makes overcoming the well-known 
  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
  the permeability of a desired molecular or atomic species, while maintaining the 
  same retention of an undesired species \cite{werber_materials_2016}.

@@

  We measured the intensity of R-pores based on the intensity of the
  d\textsubscript{100} peak (the leading peak closest to $q_r$ = 0) of the $q_r$
  cross-section of the patterns at $q_z$=0 (Figure~\ref{fig:saxs_xsection}). The
  beamstop covers most of the small angle reflections in the experimental 2D WAXS
  pattern. In order to compare the simulated intensity of R-pores to experiment,
  we used the $q_r$ cross-section of 2D SAXS which was generated from the same
  sample (Figure S-TBD). Since the d\textsubscript{110} peak is
  %BJC12: verify which crystal plane it is.  
  partially exposed in the experimental 2D WAXS pattern, we normalized the
  experimental 2D SAXS cross-section by matching the intensity of the
  d\textsubscript{110} peak between it and the experimental 2D WAXS
  cross-section. It is possible that the peak is not fully captured in the 2D
  WAXS pattern and that we have underestimated the intensity of R-pores in the
  experimental pattern.

@@

  When two vertically stacked monomer head groups hydrogen bond with a shared
  water molecule, the monomers are drawn closer together (as illustrated in
  Figure~\ref{fig:hbond_visualization}), which creates an asymmetry that allows
  R-double to appear. If a monomer head group shares a hydrogen bonded water
  molecule with a head group above itself, it will be less likely to share a
  water molecule with a head group below it due to geometric constraints. The
  monomer head group below can just as easily share a water molecule with a head
  group below itself. In this scenario, the centers of each pair are 2 times the
  $\pi$-stacking distance apart which would lead to R-double (much like the
  configuration in Figure~\ref{fig:staggered_rzplot_norestraints}). There are a
  modest number of occurrences of this scenario, which we quantify in further
  detail in the Supporting Information, Section S-TBD.

>>>>

  Nanostructured porous membranes made from the cross-linked hexagonal phase
  of self-assembled lyotropic liquid crystals (LLCs) are a promising material
  for selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular,
  we compare simulated X-ray diffraction (XRD) patterns with experimental XRD
  to quantify and understand the differences between simulation and experiment.
  We find that pores are likely composed of 5 columns of stacked LLC monomers
  which surround each hydrophilic core. We learned that these columns likely 
  move independently of each other over time scales which we can not simulate.
  We find that some of the structure previously attributed to monomer tail 
  tilt is likely instead due to ordered tail packing. Although this system
  has been reported as dry, we show that small amounts of water are necessary
  to fully reproduce all features from the experimental XRD pattern due to
  asymmetries introduced by hydrogen bonds between monomer head groups and
  water molecules. We explore the composition and structure of the nanopores
  and reveal that there exists a composition gradient rather than an abrupt
  partition between the hydrophilic and hydrophobic regions. 
  % BJC13: If this is to be included here, it needs some rewording. I allude to
  % this when talking about the independently moving columns above. 
  %A caveat is
  %that the time scales of the dynamics are extremely long for this system, 
  %resulting in simulated structures that appear too ordered which requires
  %careful examination of the metastable states observed in order to draw any
  %conclusions. 
  The clearer picture of the nanoscopic structure of these membranes provided
  in this study will enable a better understanding of the mechanisms of small
  molecule transport within these nanopores.

  \end{abstract}

@@

  More highly selective membranes would be extremely useful for both the
  recovery of valuable dissolved species from and the purification of complex
  aqueous solutions. The supply of fresh water is being increasingly strained
  by population growth, industrialization and climate change
  \cite{elimelech_future_2011}. Consequently, people are turning towards
  treatment of unconventional water sources such as seawater and wastewater. 
  In addition to obtaining fresh water from them, these water sources, 
  particularly municipal and industrial wastewater, have been found to contain 
  valuable resources which can be separated using highly selective membranes~\cite{sales_resource_2015}.

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane 
  filtration processes that can be used to separate solutes on the order of 1 
  nm in size and smaller, as well as ions. Both apply hydraulic pressure to 
  the feed solution in order to overcome osmotic pressure and force water
  and unfiltered components through the membrane. RO membranes typically have
  a thin film composite architecture which, at minimum, consists of a porous
  support layer and a dense polymer active layer. Separation occurs based on the differences in
  solubility and diffusivity of feed solution constituents with respect to the
  active layer. Nanofiltration membranes have explicit pores on the order of 1
  nm in diameter which are typically tortuous and polydisperse in size. They 
  separate primarily based on size exclusion as well as Donnan exclusion if the
  membrane has a surface charge~\cite{van_der_bruggen_review_2003}.

  The energy consumed during the purification of seawater and wastewater 
  using RO and NF membranes can be reduced by increasing their selectivity 
  towards water \cite{werber_materials_2016}. While researchers often focus on
  improving membrane permeability, the flux of water through a given membrane,
  it has been shown that increased water permeability only slightly decreases
  energy requirements \cite{cohen-tanugi_quantifying_2014}.
%  since the main energy requirement for a membrane separation is that needed
%  to create the hydraulic pressure necessary in order to overcome the 
%  osmotic pressure of the feed solution\cite{cohen-tanugi_quantifying_2014}. 
  Often times, multiple membrane passes are required in order to sufficiently
  reduce contaminant levels. For example, boric acid is toxic to humans and 
  is present in seawater at levels well above acceptable drinking standards.
  It is difficult to sufficiently separate boric acid in a single pass, meaning
  multiple passes are needed and energy requirements nearly double \cite{fritzmann_state---art_2007}.
%  This is necessary in the case where high-purity water
%  is needed \cite{singh_production_2009}, when separating small neutral solutes \cite{ozaki_rejection_2002} or
%  when removing boric acid from seawater \cite{fritzmann_state---art_2007}.  
  The most effective way to reduce energy consumption
  would be to reduce the number of membrane passes for a given system using 
  more selective membranes. 
%   More highly selective membranes have the potential
%  to reduce the number of membrane passes in addition to eliminating the need 
%  for other post-treatment steps necessitated by incomplete separations.
  %dissolved species which negatively
%  affect human health.

  Not only will increased selectivity reduce the energy consumption of membrane
  processes, but it offers a way to recover valuable resources from the feed 
  streams. Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, but also contain valuable dissolved solutes. Micropollutants include
  pharmaceutical and personal care products, hormones, pesticides and industrial
  chemicals, all of which have adverse effects on human health even at low 
  concentrations\cite{schwarzenbach_challenge_2006}. A selective membrane has the % There is a need for membranes
  potential to reject these components while still allowing the passage of valuable 
  resources for further processing and recovery.

  Municipal wastewaters are rich in carbon, nitrogen and phosphorus-containing 
  compounds. The recovery of such products, which can be achieved using a 
  selective membrane, has numerous potential uses \cite{sales_resource_2015}. 
  For example, nitrogen and phosphorus recovery can help sustain fertilizer 
  production which will help meet global food demand as population continues
  to increase \cite{xie_membrane-based_2016}.

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater\cite{werber_materials_2016}. For example, flowback water, produced
  during hydraulic fracturing of shale formations consists of relatively high
  concentrations of salts, metals, and soluble organic compounds. The majority
  of this water is disposed through deep well injection, however there is a growing public concern about
  its management which has prompted the use of separation technologies such as 
  RO and NF in order to reduce the volume of contaminated water~\cite{gregory_water_2011}.
  Some of the dissolved organic compounds in flowback water, such as acetate, 
  are potentially valuable and can be recovered with highly selective membranes
  ~\cite{dischinger_application_2017}.

%  More highly selective membranes would be extremely useful for the recovery of
%  dissolved species in complex aqueous and organic solutions. For example,
%  selective membranes can further progress forward osmosis processes by reducing
%  the cost of separating potable water from draw
%  solutions~\cite{mccutcheon_novel_2005}. Additionally, flowback water (FW)
%  produced during hydraulic fracturing is a complex wastewater full of
%  potentially valuable dissolved organic compounds such as acetate that can be
%  recovered using highly selective membranes \cite{dischinger_application_2017}.
%  Finally, a significant amount of nitrogen and phosphorus from fertilizer is
%  found in wastewater. Selective membranes offer a route towards low energy
%  nutrient recovery from these wastewaters which can help sustain fertilizer
%  production and subsequently food production into the future
%  \cite{xie_membrane-based_2016}.
  % as we near peak production of minable phosphorus rock .

@@

  Current commercial RO and NF membranes suffer limitations to their selectivity
  because their structures incorporate some degree of randomness. Although scalable, their 
  fabrication involves the spontaneous assembly of polymers into disordered structures
  which offer separation pathways that are tortuous and polydisperse in size 
  \cite{werber_materials_2016}. Tortuosity increases the effective length that a 
  solute must travel, while polydispersity limits membrane selectivity.

%BJC13: You can stop here, I haven't changed anything below.

%  This makes overcoming the well-known 
%  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
%  the permeability of a desired molecular or atomic species, while maintaining the 
%  same retention of an undesired species \cite{werber_materials_2016}.

@@

  We measured the intensity of R-pores based on the intensity of the
  d\textsubscript{100} peak (the leading peak closest to $q_r$ = 0) of the $q_r$
  cross-section of the patterns at $q_z$=0 (Figure~\ref{fig:saxs_xsection}). The
  beamstop covers most of the small angle reflections in the experimental 2D WAXS
  pattern. In order to compare the simulated intensity of R-pores to experiment,
  we used the $q_r$ cross-section of 2D SAXS which was generated from the same
  sample (Figure S-\ref{S-fig:2DSAXS}). Since the d\textsubscript{200} peak is
  partially exposed in the experimental 2D WAXS pattern, we normalized the
  experimental 2D SAXS cross-section by matching the intensity of the
  d\textsubscript{200} peak between it and the experimental 2D WAXS
  cross-section. It is possible that the peak is not fully captured in the 2D
  WAXS pattern and that we have underestimated the intensity of R-pores in the
  experimental pattern.

@@

  When two vertically stacked monomer head groups hydrogen bond with a shared
  water molecule, the monomers are drawn closer together (as illustrated in
  Figure~\ref{fig:hbond_visualization}), which creates an asymmetry that allows
  R-double to appear. If a monomer head group shares a hydrogen bonded water
  molecule with a head group above itself, it will be less likely to share a
  water molecule with a head group below it due to geometric constraints. The
  monomer head group below can just as easily share a water molecule with a head
  group below itself. In this scenario, the centers of each pair are 2 times the
  $\pi$-stacking distance apart which would lead to R-double (much like the
  configuration in Figure~\ref{fig:staggered_rzplot_norestraints}). There are a
  modest number of occurrences of this scenario, which we quantify in further
  detail in the Supporting Information, Section S-\ref{S-section:rdouble}.

### 2018-08-28 Ben_Manuscripts/structure_paper/supporting_information.tex
Updates to the introduction

>>>>

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{2DSAXS_axes.png}
  \caption{Two dimensional small angle X-ray scattering experiments show
  reflections along the $q_r$ axis at $q_z$=0. The spacing between the reflections
  is characteristic of a hexagonal phase and is directly related to the distance
  between pores. The reflections are arced because the pores are not perfectly
  aligned. A perfectly aligned system would show dots at each peak location. An
  isotropically aligned system would show concentric rings about the origin which
  pass through each peak location. The d\textsubscript{100} and d\textsubscript{110} 
  peaks (the first two pairs of reflections from origin located at $|\mathbf{q}| 
  \approx 0.18~\AA^{-1}$ and $|\mathbf{q}| \approx 0.31~\AA^{-1}$) are not visible
  in the 2D WAXS pattern. The edges of the d\textsubscript{200} reflection (the third
  pair of reflections located at $|\mathbf{q}| \approx 0.36~\AA^{-1}$) are slightly
  visible in the 2D WAXS pattern, so we used them to approximate the experimental
  intensity of R-pores.}\label{fig:2DSAXS}
  \end{figure}
